---
title: "[Tit-for-Tat] Strategy Identification and Payoff Analysis" 
author: "Hana Kwon"
date: "Oct 27, 2024"
output: 
  pdf_document:
    latex_engine: xelatex
---

### Introduction

This document details the approach used to identify player strategies *(Tit-for-Tat, Grim, Always Cooperate, and Always Defect)* and calculate payoffs in an experimental dataset.

### Definitions and Assumptions

-   **Payoff Values**

    -   `r`: *(Reward)* Payoff for mutual cooperation
    -   `s`: *(Sucker)* Payoff when one cooperates and the other defects
    -   `t`: *(Temptation)* Payoff for defection when the other cooperates
    -   `p`: *(Punishment)* Payoff for mutual defection

-   **Strategies**

    1.  **Always Cooperate**: Player cooperates in all rounds.
    2.  **Always Defect**: Player defects in all rounds.
    3.  **Grim**: Player starts by cooperating and defects permanently if the opponent defects.
    4.  **Tit-for-Tat**: Player mirrors the opponent’s previous move.

-   **Dataset Information**

    -   `session`: Experiment session number.

    -   `id`: Participant ID.

    -   `oid`: Partner's ID.

    -   `supergame`: Match number.

    -   `round`: Round number within a supergame.

    -   `horizon`: Length of supergame.

    -   `coop`: Cooperation indicator (1 if cooperated, 0 otherwise).

    -   `r`, `s`, `t`, `p`: Payoff values based on cooperation and defection as described above.

### **Analysis Overview**

-   **Core Analysis**

    1.  Data Loading and Initial Exploration

    2.  Data Preparation and Preprocessing

    3.  Payoff Analysis

        3.1 Calculate Payoff Based on Actions

        3.2 Round-by-Round Average Payoff Calculation

        3.3 Average Payoff and Variance by Player

        3.4 Calculation and Visualization of T Value

        3.5 Cross-Tabulation of T Values by Game Length

    4.  Strategy Identification

    5.  Strategy Payoff Analysis

-   **Extended Analysis**

    5.  Frequency-Weighted Payoff Calculation

    6.  Performance Against Non-Fixed Strategies

    7.  Expected Payoff Simulation for Hypothetical Tit-for-Tat Player

    8.  Visualizations and Graphical Analysis

------------------------------------------------------------------------

## **Core Analysis**

### Step 1: Data Loading, Initial Exploration, and Preparation

-   **Objective:** Load the dataset and examine its structure to ensure successful data import and check for any missing values.

```{r}
### 1.1 Load Necessary Libraries and Dataset
library(dplyr)
library(readr)
library(ggplot2)

file_path <- "~/Desktop/[Tit for Tat] Prof.MacLeod_Hana Kwon/Embrey_2018a_new_data.txt"
data <- read.table(file_path, header = TRUE, sep = "\t", stringsAsFactors = FALSE)

### 1.2 Initial Data Exploration
str(data)      # Check the structure of the data
summary(data)  # Summary of data to examine distributions and any NA values
```

### **Step 2: Data Preparation and Preprocessing**

-   **Objective:** Prepare player and opponent data frames to align cooperation values and payoff values for each round.

```{r}
### 2.1 Create Player and Opponent Data Frames
df_self <- data %>%
  select(id, oid, supergame, round, horizon, coop, r, s, t, p) %>% 
  rename(player_id = id, opponent_id = oid, player_coop = coop)

df_opp <- data %>%
  select(id, oid, supergame, round, horizon, coop) %>%  
  rename(opponent_id = id, player_id = oid, opponent_coop = coop)

### 2.2 Merge Player and Opponent Data
df_merged <- df_self %>%
  left_join(df_opp, by = c("player_id", "opponent_id", "supergame", "round", "horizon"))  

# Check structure of merged dataframe
str(df_merged) 
```

### **Step 3: Descriptive Payoff Analysis**

#### **3.1 Payoff Calculation Based on Actions**

-   **Objective:** Calculate the payoff based on cooperation and defection combinations for each round and assign values to the `payoff` column.

```{r}
df_merged <- df_merged %>%
  mutate(payoff = case_when(
    player_coop == 1 & opponent_coop == 1 ~ r,  # Both Cooperate
    player_coop == 1 & opponent_coop == 0 ~ s,  # Only Player Cooperates
    player_coop == 0 & opponent_coop == 1 ~ t,  # Only Player Defects
    player_coop == 0 & opponent_coop == 0 ~ p,  # Both Defect
    TRUE ~ NA_real_                             # Default value for any unspecified cases
  ))
```

#### 3.2 Round-by-Round Average Payoff Calculation

-   **Objective:** Calculate the round-by-round average payoff for each player and visualize changes in performance over each round.

```{r}
round_avg_payoff <- df_merged %>%
  group_by(player_id, round) %>%
  summarize(avg_round_payoff = mean(payoff, na.rm = TRUE), .groups = "drop")

print(round_avg_payoff)

# Visualization: Round-by-Round Payoff Distribution for Each Player with Grouping and Color
ggplot(round_avg_payoff, aes(x = round, y = avg_round_payoff, color = factor(player_id), group = player_id)) +
  geom_line(alpha = 0.5) +
  labs(title = "Round-by-Round Average Payoff for Each Player", x = "Round", y = "Average Payoff") +
  theme_minimal() +
  theme(legend.position = "none")
```

#### 3.3 Average Payoff and Variance Calculation per Player

-   **Objective:** Calculate the overall average payoff and payoff variance for each player to understand the distribution of player performance.

```{r}
avg_payoff_variance <- df_merged %>%
  group_by(player_id) %>%
  summarize(
    avg_payoff = mean(payoff, na.rm = TRUE),
    payoff_variance = var(payoff, na.rm = TRUE)
  )

print(avg_payoff_variance)

# Visualization: Distribution of Average Payoffs
ggplot(avg_payoff_variance, aes(x = avg_payoff)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  labs(title = "Distribution of Average Payoffs", x = "Average Payoff", y = "Count") +
  theme_minimal()
```

#### 3.4 'T' Value Calculation and Distribution Visualization

-   **Objective:** Calculate T value (the number of remaining rounds when a player first defects) and visualize the distribution of T values.

```{r}
df_merged <- df_merged %>%
  group_by(player_id, supergame) %>%
  mutate(
    first_defect_round = ifelse(player_coop == 0 & !is.na(player_coop), round, NA),
    T = ifelse(!is.na(first_defect_round), horizon - first_defect_round, ifelse(all(player_coop == 1, na.rm = TRUE), 0, NA))
  ) %>%
  ungroup() 

print(df_merged %>% select(player_id, supergame, round, player_coop, horizon, T))

# Visualization: Distribution of T Values
ggplot(df_merged %>% filter(!is.na(T)), aes(x = T)) +
  geom_histogram(bins = 30, fill = "green", color = "black") +
  labs(title = "Distribution of T Values", x = "T (Remaining Periods at First Defection)", y = "Count") +
  theme_minimal()
```

#### 3.5 Cross-Tabulation of T Values by Game Length

-   **Objective:** Summarize T values by game length (horizon) to show mean and standard deviation of T for each horizon and visualize these statistics.

```{r}
T_summary <- df_merged %>%
  group_by(horizon) %>%
  summarise(
    mean_T = mean(T, na.rm = TRUE),
    sd_T = sd(T, na.rm = TRUE),
    count = n()
  )

print(T_summary)


### Visualization: 

# Average T Value by Game Length Visualization
ggplot(T_summary, aes(x = factor(horizon), y = mean_T, fill = factor(horizon))) +
  geom_bar(stat = "identity") +
  labs(title = "Average T Value by Game Length", x = "Game Length (Horizon)", y = "Average T") +
  theme_minimal()

# Standard Deviation of T Value by Game Length Visualization
ggplot(T_summary, aes(x = factor(horizon), y = sd_T, fill = factor(horizon))) +
  geom_bar(stat = "identity") +
  labs(title = "Standard Deviation of T Value by Game Length", x = "Game Length (Horizon)", y = "Standard Deviation of T") +
  theme_minimal()

```

### **Step 4: Strategy Identification**

-   **Objective:** Define a function to classify player strategies, and then merge the identified strategies with the main dataset and calculate the average payoff and payoff variance for each strategy.

```{r}
# Load ggplot2 library for visualization
library(ggplot2)

### 4.1 Define Strategy Identification Function
identify_strategy <- function(player_coop, opponent_coop) {
  # Tit for Tat
  tit_for_tat <- if (length(player_coop) > 1) all(player_coop[-1] == lag(opponent_coop)[-1], na.rm = TRUE) else FALSE
  # Grim Strategy
  grim <- all((player_coop == 1) | (cumsum(opponent_coop == 0) > 0), na.rm = TRUE)
  # Always Cooperate
  always_cooperate <- all(player_coop == 1)
  # Always Defect
  always_defect <- all(player_coop == 0)
  
  ## Determine Strategy
  if (tit_for_tat) return("Tit for Tat")
  else if (grim) return("Grim")
  else if (always_cooperate) return("Always Cooperate")
  else if (always_defect) return("Always Defect")
  else return("Other")
}

# Apply Strategy Identification to df_merged
df_merged <- df_merged %>%
  group_by(player_id, supergame) %>%
  mutate(strategy_label = identify_strategy(player_coop, opponent_coop)) %>%
  ungroup()

# Group by strategy_label and count
strategy_counts <- df_merged %>%
  group_by(strategy_label) %>%
  summarize(count = n())

print(strategy_counts)

#==================================================================================#

# Visualize Strategy Counts
ggplot(strategy_counts, aes(x = strategy_label, y = count, fill = strategy_label)) +
  geom_bar(stat = "identity") +
  labs(title = "Strategy Counts", x = "Strategy", y = "Count") +
  theme_minimal()
```

### **Step 5: Strategy Payoff Analysis**

-   **Objective:** Merge the identified strategies with the main dataset and calculate the average payoff and payoff variance for each strategy.

```{r}

### 5.1 Apply Strategy Identification to Each Player and Supergame
df_strategies <- df_merged %>%
  group_by(player_id, supergame) %>%
  summarise(strategy = identify_strategy(player_coop, opponent_coop), .groups = 'drop')

### 5.2 Merge Identified Strategies with Original Data
df_merged <- df_merged %>%
  left_join(df_strategies, by = c("player_id", "supergame"))

### 5.3 Calculate Average Payoff by Strategy
strategy_payoff_summary <- df_merged %>%
  group_by(strategy) %>%
  summarise(
    avg_strategy_payoff = mean(payoff, na.rm = TRUE),
    payoff_variance = var(payoff, na.rm = TRUE),
    count = n()
  )

# View Results
print(strategy_payoff_summary)

#==================================================================================#

## Visualize Average Payoff by Strategy & Payoff Variance by Strategy

# Average Payoff by Strategy
ggplot(strategy_payoff_summary, aes(x = strategy, y = avg_strategy_payoff, fill = strategy)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Payoff by Strategy", x = "Strategy", y = "Average Payoff") +
  theme_minimal()

# Payoff Variance by Strategy
ggplot(strategy_payoff_summary, aes(x = strategy, y = payoff_variance, fill = strategy)) +
  geom_bar(stat = "identity") +
  labs(title = "Payoff Variance by Strategy", x = "Strategy", y = "Payoff Variance") +
  theme_minimal()
```

------------------------------------------------------------------------

## **Extended Analysis**

### Step 6: Frequency-Weighted Payoff Calculation

-   **Objective:** This calculation weights each strategy’s average payoff by its frequency, providing a clearer view of strategy effectiveness.

```{r}

### 6 Frequency-Weighted Average Payoff Calculation
weighted_payoff_summary <- strategy_payoff_summary %>%
  mutate(weighted_avg_payoff = avg_strategy_payoff * (count / sum(count)))

# Print Weighted Results
print(weighted_payoff_summary)

# Visualization
ggplot(weighted_payoff_summary, aes(x = strategy, y = weighted_avg_payoff, fill = strategy)) +
  geom_bar(stat = "identity") +
  labs(title = "Frequency-Weighted Average Payoff by Strategy", x = "Strategy", y = "Weighted Average Payoff") +
  theme_minimal()

```

### Step 7: Performance Against Non-Fixed Strategies

-   **Objective:** This analysis compares each fixed strategy’s performance when interacting with non-fixed strategies, highlighting adaptability.

```{r}

### 7 Payoff Analysis Against Non-Fixed Strategies
non_fixed_performance <- df_merged %>%
  filter(strategy != "Always Cooperate" & strategy != "Always Defect") %>%
  group_by(strategy) %>%
  summarise(
    avg_payoff_against_non_fixed = mean(payoff, na.rm = TRUE),
    variance_against_non_fixed = var(payoff, na.rm = TRUE)
  )

# Print Non-Fixed Performance Results
print(non_fixed_performance)

# Visualization
ggplot(non_fixed_performance, aes(x = strategy, y = avg_payoff_against_non_fixed, fill = strategy)) +
  geom_bar(stat = "identity") +
  labs(title = "Payoff Against Non-Fixed Strategies", x = "Strategy", y = "Average Payoff") +
  theme_minimal()
```

### Step 8: Expected Payoff Simulation for Hypothetical Tit-for-Tat Player

-   **Objective:** In this step, we estimate the expected payoff for a hypothetical Tit-for-Tat player when competing against various other strategies. This analysis provides insights into how a Tit-for-Tat strategy might perform on average against other identified strategies.

```{r}

### 8 Simulating Tit-for-Tat Performance

# Calculating Expected Payoff for Tit-for-Tat vs Opponent Strategies
hypothetical_tft_performance <- df_merged %>%
  filter(strategy == "Tit for Tat") %>%
  group_by(opponent_strategy = strategy) %>%
  summarise(
    avg_expected_payoff = mean(payoff, na.rm = TRUE),
    variance_expected_payoff = var(payoff, na.rm = TRUE),
    count = n()
  )

# View Results as a Table
print(hypothetical_tft_performance)
```

### **Step** 9**:** Visualizations and Graphical Analysis

-   **Objective:** Provides a visualization of strategy payoff distributions and strategy counts, offering a visual comparison of each strategy’s performance and popularity.

```{r}

### 9 Visualizations

library(ggplot2)
library(ggridges)

# 1. Bar Plot: Average Payoff by Strategy
# This bar plot shows the average payoff for each strategy, allowing a straightforward comparison of average performance.
ggplot(strategy_payoff_summary, aes(x = strategy, y = avg_strategy_payoff, fill = strategy)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Payoff by Strategy", x = "Strategy", y = "Average Payoff") +
  theme_minimal()

# 2. Box Plot: Payoff Distribution by Strategy
# This box plot displays the distribution of payoffs for each strategy, highlighting median, quartiles, and potential outliers.
ggplot(df_merged, aes(x = strategy, y = payoff, fill = strategy)) +
  geom_boxplot() +
  labs(title = "Payoff Distribution by Strategy", x = "Strategy", y = "Payoff") +
  theme_minimal()

# 3. Violin Plot: Density of Payoff by Strategy
# The violin plot shows the density distribution of payoffs for each strategy, combining box plot and density plot features.
ggplot(df_merged, aes(x = strategy, y = payoff, fill = strategy)) +
  geom_violin() +
  labs(title = "Density of Payoff by Strategy", x = "Strategy", y = "Payoff") +
  theme_minimal()

# 4. Density Plot: Density of Payoffs by Strategy
# This density plot presents the probability density of payoffs across strategies, allowing visual comparison of payoff distribution shapes.
ggplot(df_merged, aes(x = payoff, color = strategy)) +
  geom_density() +
  labs(title = "Density Plot of Payoffs by Strategy", x = "Payoff", y = "Density") +
  theme_minimal()

# 5. Ridgeline Plot: Payoffs by Strategy
# The ridgeline plot provides a layered view of payoff distributions by strategy, making it easier to compare their overall distributions.
ggplot(df_merged, aes(x = payoff, y = strategy, fill = strategy)) +
  geom_density_ridges(bandwidth = 1.75) +
  labs(title = "Ridgeline Plot of Payoffs by Strategy", x = "Payoff", y = "Strategy") +
  theme_minimal()

# 6. Scatter Plot: Payoffs by Strategy
# This scatter plot shows individual payoff points for each strategy, adding some noise with width to reveal overlapping data points.
ggplot(df_merged, aes(x = strategy, y = payoff, color = strategy)) +
  geom_jitter(width = 0.2) +
  labs(title = "Scatter Plot of Payoffs by Strategy", x = "Strategy", y = "Payoff") +
  theme_minimal()
```
