---
title: "TFT_241109_combined_modules"
author: "Hana Kwon, WB Macleod"
date: "Nov 9, 2024"
output:
  pdf_document:
    latex_engine: xelatex
---

# Introduction

This document details the approach used to identify player strategies *(Tit-for-Tat, Grim, Always Cooperate, and Always Defect)* and calculate payoffs in an experimental dataset.

## Definitions and Assumptions

-   **Payoff Values**

    -   `r`: *(Reward)* Payoff for mutual cooperation
    -   `s`: *(Sucker)* Payoff when one cooperates and the other defects
    -   `t`: *(Temptation)* Payoff for defection when the other cooperates
    -   `p`: *(Punishment)* Payoff for mutual defection

-   **Strategies**

    1.  **Always Cooperate**: Player cooperates in all rounds.
    2.  **Always Defect**: Player defects in all rounds.
    3.  **Grim**: Player starts by cooperating and defects permanently if the opponent defects.
    4.  **Tit-for-Tat**: Player mirrors the opponent’s previous move.

-   **Dataset Information**

    -   `session`: Experiment session number.

    -   `id`: Participant ID.

    -   `oid`: Partner's ID.

    -   `supergame`: Match number.

    -   `round`: Round number within a supergame.

    -   `horizon`: Length of supergame.

    -   `coop`: Cooperation indicator (1 if cooperated, 0 otherwise).

    -   `r`, `s`, `t`, `p`: Payoff values based on cooperation and defection as described above.

## **Analysis Overview**

-   **Core Analysis**

    1.  Data Loading and Initial Exploration

    2.  Data Preparation and Preprocessing

    3.  Payoff Analysis

        3.1 Calculate Payoff Based on Actions

        3.2 Round-by-Round Average Payoff Calculation

        3.3 Average Payoff and Variance by Player

        3.4 Calculation and Visualization of T Value

        3.5 Cross-Tabulation of T Values by Game Length

    4.  Strategy Identification

    5.  Strategy Payoff Analysis

-   **Extended Analysis**

    5.  Frequency-Weighted Payoff Calculation

    6.  Performance Against Non-Fixed Strategies

    7.  Expected Payoff Simulation for Hypothetical Tit-for-Tat Player

    8.  Visualizations and Graphical Analysis

------------------------------------------------------------------------

# **Data Preparation**

### Step 1: Data Loading, Initial Exploration, and Preparation

-   **Objective:** Load the dataset and examine its structure to ensure successful data import and check for any missing values.

    ```{r}
    ### 1.1 Load Necessary Libraries and Dataset
    library(dplyr)
    library(readr)
    library(ggplot2)

    file_data <- "/Users/hanakwon/Desktop/0_RA_McLeod_TFT/Embrey_2018a_new_data.txt"
    data <- read.table(file_data, header = TRUE, sep = "\t", stringsAsFactors = FALSE)

    ### 1.2 Initial Data Exploration
    str(data)      # Check the structure of the data
    summary(data)  # Summary of data to examine distributions and any NA values

    ```

------------------------------------------------------------------------

### **Step 2:** Data Preparation, Preprocessing, and PD Difficulty Setting

-   **Objective:** Prepare player and opponent data frames to align cooperation values and payoff values for each round. Classify games by PD difficulty (EasyPD or HardPD) based on payoff values.

-   **Code Updates (Nov 1, 2024) =\> 2x2 Design**

    ```{r}
    ###Updated 
    # Set normalized payoff values
    data <- data %>%
      mutate(
        r = 1.0,              # Reward for mutual cooperation
        s = -l,               # Sucker's payoff
        t = 1.0 + g,          # Temptation payoff
        p = 0.0               # Punishment for mutual defection
      )

    # Create player and opponent data frames
    df_self <- data %>%
      select(id, oid, supergame, round, horizon, coop, r, s, t, p) %>%
      rename(player_id = id, opponent_id = oid, player_coop = coop)

    df_opp <- data %>%
      select(id, oid, supergame, round, horizon, coop) %>%
      rename(opponent_id = id, player_id = oid, opponent_coop = coop)

    # Merge player and opponent data
    df_merged <- df_self %>%
      left_join(df_opp, by = c("player_id", "opponent_id", "supergame", "round", "horizon"))


    # Add PD difficulty column
    df_merged <- df_merged %>%
      mutate(pd_difficulty = case_when(
        abs(t - 2) < 0.1 & abs(s + 1.41) < 0.1 ~ "EasyPD",
        abs(t - 4) < 0.1 & abs(s + 2.8) < 0.1 ~ "HardPD",
        TRUE ~ NA_character_
      ))
    ```

------------------------------------------------------------------------

# Core Analysis

### **Step 3: Descriptive Payoff Analysis**

#### **3.1 Payoff Calculation Based on Actions**

-   **Objective:** Calculate the payoff based on cooperation and defection combinations for each round and assign values to the `payoff` column.

```{r}

#Updated

# Calculate the Payoff Column Based on Actions
df_merged <- df_merged %>%
  mutate(payoff = case_when(
    player_coop == 1 & opponent_coop == 1 ~ r,  # Both Cooperate
    player_coop == 1 & opponent_coop == 0 ~ s,  # Only Player Cooperates
    player_coop == 0 & opponent_coop == 1 ~ t,  # Only Player Defects
    player_coop == 0 & opponent_coop == 0 ~ p,  # Both Defect
    TRUE ~ NA_real_                             # Default value for any unspecified cases
  ))

# Calculate payoffs
df_merged <- df_merged %>%
  mutate(payoff = case_when(
    player_coop == 1 & opponent_coop == 1 ~ r,  
    player_coop == 1 & opponent_coop == 0 ~ s,  
    player_coop == 0 & opponent_coop == 1 ~ t,  
    player_coop == 0 & opponent_coop == 0 ~ p,  
    TRUE ~ NA_real_
  ))

# Verify that the 'payoff' column exists
if (!"payoff" %in% colnames(df_merged)) {
  stop("Error: 'payoff' column was not created.")
} else {
  print("Payoff column created successfully.")
}

```

#### 3.2 Round-by-Round Average Payoff Calculation

-   **Objective:** Calculate the round-by-round average payoff for each player and visualize changes in performance over each round.

-   **Code Updates (Nov 1, 2024) =\> 2x2 Design**

    ```{r}
    # 3.2 Round-by-Round Average Payoff Calculation with 2x2 Design

    round_avg_payoff <- df_merged %>%
      group_by(player_id, round, horizon, pd_difficulty) %>%
      summarize(avg_round_payoff = mean(payoff, na.rm = TRUE), .groups = "drop")

    print(round_avg_payoff)

    # Visualization: Round-by-Round Payoff Distribution for Each Player in 2x2 Design
    ggplot(round_avg_payoff, aes(x = round, y = avg_round_payoff, color = factor(player_id), group = player_id)) +
      geom_line(alpha = 0.5) +
      labs(title = "Round-by-Round Average Payoff for Each Player", x = "Round", y = "Average Payoff") +
      facet_grid(pd_difficulty ~ horizon) +  # 2x2 design: Difficulty by Game Length
      theme_minimal() +
      theme(legend.position = "none")

    ```

#### 3.3 Average Payoff and Variance Calculation per Player

-   **Objective:** Calculate the overall average payoff and payoff variance for each player to understand the distribution of player performance.

-   **Code Updates (Nov 1, 2024) =\> 2x2 Design**

    ```{r}
    avg_payoff_variance <- df_merged %>%
      group_by(player_id, horizon, pd_difficulty) %>%
      summarize(
        avg_payoff = mean(payoff, na.rm = TRUE),
        payoff_variance = var(payoff, na.rm = TRUE),
        .groups = "drop"
      )

    # Visualization with 2x2 Design
    ggplot(avg_payoff_variance, aes(x = avg_payoff)) +
      geom_histogram(bins = 30, fill = "blue", color = "black") +
      facet_grid(pd_difficulty ~ horizon) +
      labs(title = "Distribution of Average Payoffs by Game Type", x = "Average Payoff", y = "Count") +
      theme_minimal()


    ```

#### 3.4 'T' Value Calculation and Distribution Visualization

-   **Objective:** Calculate T value (the number of remaining rounds when a player first defects) and visualize the distribution of T values.

-   **Code Updates (Nov 1, 2024) =\> 2x2 Design**

    ```{r}

    df_merged <- df_merged %>%
      group_by(player_id, supergame) %>%
      mutate(
        first_defect_round = ifelse(player_coop == 0 & !is.na(player_coop), round, NA),
        T = ifelse(!is.na(first_defect_round), horizon - first_defect_round, ifelse(all(player_coop == 1, na.rm = TRUE), 0, NA))
      ) %>%
      ungroup() 

    print(df_merged %>% select(player_id, supergame, round, player_coop, horizon, T))


    # Visualization: Distribution of T Values with 2x2 Design
    ggplot(df_merged %>% filter(!is.na(T)), aes(x = T)) +
      geom_histogram(bins = 30, fill = "green", color = "black") +
      facet_grid(pd_difficulty ~ horizon) +
      labs(title = "Distribution of T Values by Game Type", x = "T (Remaining Periods at First Defection)", y = "Count") +
      theme_minimal()
    ```

#### 3.5 Cross-Tabulation of T Values by Game Length

-   **Objective:** Summarize T values by game length (horizon) to show mean and standard deviation of T for each horizon and visualize these statistics.

-   **Code Updates (Nov 1, 2024) =\> 2x2 Design**

    ```{r}
    # Update T_summary to group by both horizon and pd_difficulty
    T_summary <- df_merged %>%
      group_by(horizon, pd_difficulty) %>%
      summarise(
        mean_T = mean(T, na.rm = TRUE),
        sd_T = sd(T, na.rm = TRUE),
        count = n(),
        .groups = "drop"
      )

    print(T_summary)

    # Visualization: Average T Value by Game Length and Difficulty
    ggplot(T_summary, aes(x = factor(horizon), y = mean_T, fill = pd_difficulty)) +
      geom_bar(stat = "identity", position = "dodge") +
      facet_wrap(~ pd_difficulty) +
      labs(title = "Average T Value by Game Length and Difficulty", x = "Game Length (Horizon)", y = "Average T") +
      theme_minimal()

    # Visualization: Standard Deviation of T Value by Game Length and Difficulty
    ggplot(T_summary, aes(x = factor(horizon), y = sd_T, fill = pd_difficulty)) +
      geom_bar(stat = "identity", position = "dodge") +
      facet_wrap(~ pd_difficulty) +
      labs(title = "Standard Deviation of T Value by Game Length and Difficulty", x = "Game Length (Horizon)", y = "Standard Deviation of T") +
      theme_minimal()


    ```

------------------------------------------------------------------------

### **Step 4: Strategy Identification**

-   **Objective:** Define a function to classify player strategies (Always Cooperate, Always Defect, Grim, Tit-for-Tat, and Experimenter) and merge the identified strategies with the main dataset.

-   **Code Updates (Nov 1, 2024) =\> 2x2 Design**

-   **Code Updates (Nov 5, 2024) =\> Added Grim strategy refinement with tolerance levels (G_0, G_1, G_2) and implemented interaction-based strategy classification.**

    ```{r}


    # 4.1: Apply Grim strategy tolerance refinement (G_0, G_1, G_2)
    df_merged <- df_merged %>%
      mutate(
        G_0 = if_else(player_coop == 1 & lag(opponent_coop, default = 1) == 0, 1, 0),
        G_1 = if_else(lag(player_coop, 1, default = 1) == 1 & lag(opponent_coop, 1, default = 1) == 0, 1, 0),
        G_2 = if_else(lag(player_coop, 2, default = 1) == 1 & lag(opponent_coop, 2, default = 1) == 0, 1, 0)
      )

    # 4.2: Define function for interaction-based strategy classification
    identify_strategy <- function(player_coop, opponent_coop, tolerance) {
      
      # Check for Always Cooperate (AC)
      if (all(player_coop == 1, na.rm = TRUE)) {
        return("Always Cooperate")
      }
      
      # Check for Always Defect (AD)
      if (all(player_coop == 0, na.rm = TRUE)) {
        return("Always Defect")
      }
      
      # Check for Grim strategy with varying levels of tolerance
      if (all((player_coop == 1) | (cumsum(opponent_coop == 0) <= tolerance), na.rm = TRUE)) {
        return(paste("Grim with Tolerance", tolerance))
      }
      
      # Check for Tit-for-Tat (TFT) - mirrors opponent's previous action
      if (length(player_coop) > 1 && all(player_coop[-1] == lag(opponent_coop)[-1], na.rm = TRUE)) {
        return("Tit for Tat")
      }
      
      # Default to Experimenters if no match
      return("Experimenters")
    }

    # 4.3: Apply the classification across interactions
    df_merged <- df_merged %>%
      group_by(player_id, opponent_id, supergame) %>%
      mutate(
        strategy_label = identify_strategy(player_coop, opponent_coop, tolerance = 0), # Strict Grim
        strategy_label_tolerance_1 = identify_strategy(player_coop, opponent_coop, tolerance = 1), # Grim T1
        strategy_label_tolerance_2 = identify_strategy(player_coop, opponent_coop, tolerance = 2) # Grim T2
      ) %>%
      ungroup()

    # 4.4: Check the consistency of strategy labels across different Grim tolerance levels
    df_merged <- df_merged %>%
      mutate(
        strategy_final = case_when(
          strategy_label == "Always Cooperate" ~ "Always Cooperate",
          strategy_label == "Always Defect" ~ "Always Defect",
          strategy_label == "Tit for Tat" ~ "Tit for Tat",
          strategy_label == "Experimenters" ~ "Experimenters",
          TRUE ~ strategy_label # Defaulting to the primary Grim strategy identified
        )
      )

    # 4.5: Print a summary of strategies by game for inspection
    strategy_summary <- df_merged %>%
      group_by(player_id, strategy_final) %>%
      summarise(
        avg_payoff = mean(payoff, na.rm = TRUE),
        payoff_variance = var(payoff, na.rm = TRUE),
        count = n(),
        .groups = "drop"
      )

    # Display strategy summary
    print(strategy_summary)

    # Visualization for average payoff by strategy
    ggplot(strategy_summary, aes(x = strategy_final, y = avg_payoff, fill = strategy_final)) +
      geom_bar(stat = "identity", position = position_dodge()) +
      labs(title = "Average Payoff by Identified Strategy", x = "Strategy", y = "Average Payoff") +
      theme_minimal()
    ```

    \

    ```{r}

    # Calculate average payoff by strategy, horizon, and PD difficulty
    strategy_payoff_difficulty <- df_merged %>%
      group_by(strategy_label, horizon, pd_difficulty) %>%
      summarise(
        avg_payoff = mean(payoff, na.rm = TRUE),
        payoff_variance = var(payoff, na.rm = TRUE),
        count = n(),
        .groups = "drop"
      )
    # Print summary table
    print(strategy_payoff_difficulty)


    #####========Updated_241101===================================================#####

    # Visualization: Average Payoff by Strategy, Horizon, and PD Difficulty with adjusted text size
    ggplot(strategy_payoff_difficulty, aes(x = strategy_label, y = avg_payoff, fill = pd_difficulty)) +
      geom_bar(stat = "identity", position = position_dodge()) +
      facet_wrap(~ horizon, scales = "free") +
      scale_fill_manual(values = c("EasyPD" = "#1f77b4", "HardPD" = "#ff7f0e", "NA" = "grey70")) +
      labs(title = "Average Payoff by Strategy, Horizon, and PD Difficulty",
           x = "Strategy",
           y = "Average Payoff",
           fill = "PD Difficulty") +
      theme_minimal() +
      theme(
        text = element_text(size = 10),  # Decrease overall text size
        axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # Rotate and reduce x-axis labels
        strip.text = element_text(size = 9),  # Decrease size of facet labels
        plot.title = element_text(size = 12, face = "bold")  # Title size adjustment
      )


    #============================================================================#
    ```

------------------------------------------------------------------------

### **Step 5: Strategy Payoff Analysis**

-   **Objective:** Analyze the average payoff and payoff variance for each identified strategy, segmented by game length (horizon) and PD difficulty level (EasyPD or HardPD). This step provides insights into the effectiveness of each strategy in different game conditions and helps identify which strategies yield higher payoffs across varying difficulty levels and game lengths.

-   **Code Updates (Nov 1, 2024) =\> 2x2 Design**

    ```{r}
    # Calculate average payoff by strategy, horizon, and PD difficulty
    strategy_payoff_summary <- df_merged %>%
      group_by(strategy_label, horizon, pd_difficulty) %>%
      summarise(
        avg_strategy_payoff = mean(payoff, na.rm = TRUE),
        payoff_variance = var(payoff, na.rm = TRUE),
        count = n(),
        .groups = "drop"
      )

    # 2x2 Design Visualization with adjusted text size
    ggplot(strategy_payoff_summary, aes(x = strategy_label, y = avg_strategy_payoff, fill = pd_difficulty)) +
      geom_bar(stat = "identity", position = position_dodge()) +
      facet_grid(pd_difficulty ~ horizon, scales = "free") +
      scale_fill_manual(values = c("EasyPD" = "#1f77b4", "HardPD" = "#ff7f0e", "NA" = "grey70")) +
      labs(title = "2x2 Design: Average Payoff by Strategy, Horizon, and PD Difficulty",
           x = "Strategy",
           y = "Average Payoff",
           fill = "PD Difficulty") +
      theme_minimal() +
      theme(
        text = element_text(size = 10),  # Adjust overall text size
        axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # Rotate and resize x-axis labels
        strip.text = element_text(size = 9),  # Decrease facet label size
        plot.title = element_text(size = 12, face = "bold")  # Adjust title size and make bold
      )


    ```

------------------------------------------------------------------------

# **Extended Analysis**

# 

### Step 6: Frequency-Weighted Payoff Calculation

-   **Objective:** This calculation weights each strategy’s average payoff by its frequency, providing a clearer view of strategy effectiveness.

-   **Code Updates (Nov 1, 2024) =\> 2x2 Design**

    ```{r}

    # Calculate Frequency-Weighted Payoff 
    weighted_payoff_summary <- strategy_payoff_summary %>%
      group_by(strategy_label, horizon, pd_difficulty) %>%
      mutate(weighted_avg_payoff = avg_strategy_payoff * (count / sum(count)))

    # Visualization with 2x2 Design and adjusted text size
    ggplot(weighted_payoff_summary, aes(x = strategy_label, y = weighted_avg_payoff, fill = strategy_label)) +
      geom_bar(stat = "identity", position = "dodge") +
      facet_grid(pd_difficulty ~ horizon, scales = "free") +
      labs(title = "Frequency-Weighted Average Payoff by Strategy, Horizon, and Difficulty",
           x = "Strategy", y = "Weighted Average Payoff") +
      theme_minimal() +
      theme(
        text = element_text(size = 10),  # Set overall text size
        axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # Rotate and resize x-axis labels
        strip.text = element_text(size = 9),  # Decrease facet label size
        plot.title = element_text(size = 12, face = "bold")  # Bold and adjust title size
      )

    ```

------------------------------------------------------------------------

### Step 7: Performance Against Non-Fixed Strategies

-   **Objective:** This analysis compares each fixed strategy’s performance when interacting with non-fixed strategies, highlighting adaptability.

<!-- -->

-   **Code Updates (Nov 8, 2024)**

#### 7.1. Approach 1. General Comparison Against Non-Fixed Strategies

In this first approach, we calculate the average payoff of each strategy when competing against any non-fixed strategy. This provides an overall measure of performance without distinguishing specific opponent strategies.

```{r}

### Approach 1: Overall Performance Against Non-Fixed Strategies

non_fixed_performance_general <- df_merged %>%
  filter(strategy_label != "Always Cooperate" & strategy_label != "Always Defect") %>%
  group_by(strategy_label) %>%
  summarise(
    avg_payoff_against_non_fixed = mean(payoff, na.rm = TRUE),
    variance_against_non_fixed = var(payoff, na.rm = TRUE)
  )

# Print General Non-Fixed Performance Results
print(non_fixed_performance_general)

# Visualization for General Non-Fixed Strategy Performance
ggplot(non_fixed_performance_general, aes(x = strategy_label, y = avg_payoff_against_non_fixed, fill = strategy_label)) +
  geom_bar(stat = "identity") +
  labs(title = "Performance Against Non-Fixed Strategies (General)", x = "Strategy", y = "Average Payoff") +
  theme_minimal()
```

#### **7.2. Approach 2:** Specific Comparison Against Each Non-Fixed Strategy

In this second approach, we expand the analysis by considering each non-fixed strategy (e.g., Tit-for-Tat, Experimenters) separately. This allows us to see how each strategy performs against specific non-fixed strategies, offering a more granular view of strategy adaptability.

```{r}

non_fixed_performance_detailed <- df_merged %>%
  filter(strategy_label %in% c("Tit for Tat", "Experimenters")) %>%
  group_by(player_strategy = strategy_label, opponent_strategy = strategy_label) %>%
  summarise(
    avg_payoff_against_non_fixed = mean(payoff, na.rm = TRUE),
    variance_against_non_fixed = var(payoff, na.rm = TRUE),
    count = n(),
    .groups = "drop"
  )

# Print Detailed Non-Fixed Performance Results
print(non_fixed_performance_detailed)

# Visualization for Detailed Non-Fixed Strategy Performance
ggplot(non_fixed_performance_detailed, aes(x = player_strategy, y = avg_payoff_against_non_fixed, fill = opponent_strategy)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title = "Performance Against Non-Fixed Strategies (Detailed)", x = "Player Strategy", y = "Average Payoff Against Non-Fixed") +
  theme_minimal()
```

#### **7.3. Additional Analysis: Performance by Specific Opponent Strategy**

This final analysis adds one more layer of specificity by grouping by both the player’s and opponent’s strategy. This allows us to see exactly how each non-fixed strategy (like Tit-for-Tat or Experimenters) performs against specific opponent strategies, rather than treating all non-fixed strategies as a single group.

```{r}
# Step 1: Add Opponent Strategy
df_merged <- df_merged %>%
  group_by(opponent_id, supergame) %>%
  mutate(opponent_strategy = first(strategy_label)) %>%
  ungroup()

# Step 2: Performance Analysis by Specific Opponent Strategy
non_fixed_performance_by_opponent <- df_merged %>%
  filter(strategy_label %in% c("Tit for Tat", "Experimenters")) %>%
  group_by(player_strategy = strategy_label, opponent_strategy) %>%
  summarise(
    avg_payoff_against_non_fixed = mean(payoff, na.rm = TRUE),
    variance_against_non_fixed = var(payoff, na.rm = TRUE),
    count = n(),
    .groups = "drop"
  )

# Print Non-Fixed Performance by Specific Opponent Strategy
print(non_fixed_performance_by_opponent)

# Visualization for Performance by Specific Opponent Strategy
ggplot(non_fixed_performance_by_opponent, aes(x = player_strategy, y = avg_payoff_against_non_fixed, fill = opponent_strategy)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(title = "Performance by Specific Opponent Strategy", x = "Player Strategy", y = "Average Payoff Against Specific Non-Fixed Strategies") +
  theme_minimal()
```

------------------------------------------------------------------------

### Step 8: Expected Payoff Simulation for Hypothetical Tit-for-Tat Player

-   **Objective:** In this step, we estimate the expected payoff for a hypothetical Tit-for-Tat player when competing against various other strategies. This analysis provides insights into how a Tit-for-Tat strategy might perform on average against other identified strategies.

-   **Code Updates (Nov 1, 2024) =\> 2x2 Design**

    ```{r}


    ### 8 Simulating Tit-for-Tat Performance

    # Calculating Expected Payoff for Tit-for-Tat vs Opponent Strategies
    hypothetical_tft_performance <- df_merged %>%
      filter(strategy_label == "Tit for Tat") %>%
      group_by(opponent_strategy = strategy_label, horizon, pd_difficulty) %>%
      summarise(
        avg_expected_payoff = mean(payoff, na.rm = TRUE),
        variance_expected_payoff = var(payoff, na.rm = TRUE),
        count = n(),
        .groups = "drop"
      )

    # Visualization with 2x2 Design
    ggplot(hypothetical_tft_performance, aes(x = opponent_strategy, y = avg_expected_payoff, fill = opponent_strategy)) +
      geom_bar(stat = "identity", position = "dodge") +
      facet_grid(pd_difficulty ~ horizon) +
      labs(title = "Expected Payoff for Tit-for-Tat Player by Game Type",
           x = "Opponent Strategy", y = "Average Expected Payoff") +
      theme_minimal()

    ```

------------------------------------------------------------------------

### **Step** 9**:** Visualizations and Graphical Analysis

-   **Objective:** Provides a visualization of strategy payoff distributions and strategy counts, offering a visual comparison of each strategy’s performance and popularity.
-   **Code Updates (Nov 1, 2024)**

```{r}

### 9 Visualizations

library(ggplot2)
library(ggridges)

# 1. Bar Plot: Average Payoff by Strategy
ggplot(strategy_payoff_summary, aes(x = strategy_label, y = avg_strategy_payoff, fill = strategy_label)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Payoff by Strategy", x = "Strategy", y = "Average Payoff") +
  theme_minimal()

# 2. Box Plot: Payoff Distribution by Strategy
ggplot(df_merged, aes(x = strategy_label, y = payoff, fill = strategy_label)) +
  geom_boxplot() +
  labs(title = "Payoff Distribution by Strategy", x = "Strategy", y = "Payoff") +
  theme_minimal()

# 3. Violin Plot: Density of Payoff by Strategy
ggplot(df_merged, aes(x = strategy_label, y = payoff, fill = strategy_label)) +
  geom_violin() +
  labs(title = "Density of Payoff by Strategy", x = "Strategy", y = "Payoff") +
  theme_minimal()

# 4. Density Plot: Density of Payoffs by Strategy
ggplot(df_merged, aes(x = payoff, color = strategy_label)) +
  geom_density() +
  labs(title = "Density Plot of Payoffs by Strategy", x = "Payoff", y = "Density") +
  theme_minimal()

# 5. Ridgeline Plot: Payoffs by Strategy
ggplot(df_merged, aes(x = payoff, y = strategy_label, fill = strategy_label)) +
  geom_density_ridges(bandwidth = 1.75) +
  labs(title = "Ridgeline Plot of Payoffs by Strategy", x = "Payoff", y = "Strategy") +
  theme_minimal()

# 6. Scatter Plot: Payoffs by Strategy
ggplot(df_merged, aes(x = strategy_label, y = payoff, color = strategy_label)) +
  geom_jitter(width = 0.2) +
  labs(title = "Scatter Plot of Payoffs by Strategy", x = "Strategy", y = "Payoff") +
  theme_minimal()

```
