---
title: "TFT_241109_module1"
author: "Hana Kwon, WB Macleod"
date: "Nov 9, 2024"
output:
  pdf_document:
    latex_engine: xelatex
---

### 

# **Data Preparation**

### Step 1: Data Loading, Initial Exploration, and Preparation

-   **Objective:** Load the dataset and examine its structure to ensure successful data import and check for any missing values.

    ```{r}
    ### 1.1 Load Necessary Libraries and Dataset
    library(dplyr)
    library(readr)
    library(ggplot2)

    file_data <- "/Users/hanakwon/Desktop/0_RA_McLeod_TFT/Embrey_2018a_new_data.txt"
    data <- read.table(file_data, header = TRUE, sep = "\t", stringsAsFactors = FALSE)

    ### 1.2 Initial Data Exploration
    str(data)      # Check the structure of the data
    summary(data)  # Summary of data to examine distributions and any NA values

    ```

------------------------------------------------------------------------

### **Step 2:** Data Preparation, Preprocessing, and PD Difficulty Setting

-   **Objective:** Prepare player and opponent data frames to align cooperation values and payoff values for each round. Classify games by PD difficulty (EasyPD or HardPD) based on payoff values.

-   **Code Updates (Nov 1, 2024) =\> 2x2 Design**

    ```{r}
    ###Updated 
    # Set normalized payoff values
    data <- data %>%
      mutate(
        r = 1.0,              # Reward for mutual cooperation
        s = -l,               # Sucker's payoff
        t = 1.0 + g,          # Temptation payoff
        p = 0.0               # Punishment for mutual defection
      )

    # Create player and opponent data frames
    df_self <- data %>%
      select(id, oid, supergame, round, horizon, coop, r, s, t, p) %>%
      rename(player_id = id, opponent_id = oid, player_coop = coop)

    df_opp <- data %>%
      select(id, oid, supergame, round, horizon, coop) %>%
      rename(opponent_id = id, player_id = oid, opponent_coop = coop)

    # Merge player and opponent data
    df_merged <- df_self %>%
      left_join(df_opp, by = c("player_id", "opponent_id", "supergame", "round", "horizon"))


    # Add PD difficulty column
    df_merged <- df_merged %>%
      mutate(pd_difficulty = case_when(
        abs(t - 2) < 0.1 & abs(s + 1.41) < 0.1 ~ "EasyPD",
        abs(t - 4) < 0.1 & abs(s + 2.8) < 0.1 ~ "HardPD",
        TRUE ~ NA_character_
      ))
    ```
