---
title: "TFT_241109_module2"
author: "Hana Kwon, WB Macleod"
date: "Nov 9, 2024"
output:
  pdf_document:
    latex_engine: xelatex
---

### 

# **Core Analysis**

### **Step 3: Descriptive Payoff Analysis**

#### **3.1 Payoff Calculation Based on Actions**

-   **Objective:** Calculate the payoff based on cooperation and defection combinations for each round and assign values to the `payoff` column.

```{r}

#Updated

# Calculate the Payoff Column Based on Actions
df_merged <- df_merged %>%
  mutate(payoff = case_when(
    player_coop == 1 & opponent_coop == 1 ~ r,  # Both Cooperate
    player_coop == 1 & opponent_coop == 0 ~ s,  # Only Player Cooperates
    player_coop == 0 & opponent_coop == 1 ~ t,  # Only Player Defects
    player_coop == 0 & opponent_coop == 0 ~ p,  # Both Defect
    TRUE ~ NA_real_                             # Default value for any unspecified cases
  ))

# Calculate payoffs
df_merged <- df_merged %>%
  mutate(payoff = case_when(
    player_coop == 1 & opponent_coop == 1 ~ r,  
    player_coop == 1 & opponent_coop == 0 ~ s,  
    player_coop == 0 & opponent_coop == 1 ~ t,  
    player_coop == 0 & opponent_coop == 0 ~ p,  
    TRUE ~ NA_real_
  ))

# Verify that the 'payoff' column exists
if (!"payoff" %in% colnames(df_merged)) {
  stop("Error: 'payoff' column was not created.")
} else {
  print("Payoff column created successfully.")
}

```

#### 3.2 Round-by-Round Average Payoff Calculation

-   **Objective:** Calculate the round-by-round average payoff for each player and visualize changes in performance over each round.

-   **Code Updates (Nov 1, 2024) =\> 2x2 Design**

    ```{r}
    # 3.2 Round-by-Round Average Payoff Calculation with 2x2 Design

    round_avg_payoff <- df_merged %>%
      group_by(player_id, round, horizon, pd_difficulty) %>%
      summarize(avg_round_payoff = mean(payoff, na.rm = TRUE), .groups = "drop")

    print(round_avg_payoff)

    # Visualization: Round-by-Round Payoff Distribution for Each Player in 2x2 Design
    ggplot(round_avg_payoff, aes(x = round, y = avg_round_payoff, color = factor(player_id), group = player_id)) +
      geom_line(alpha = 0.5) +
      labs(title = "Round-by-Round Average Payoff for Each Player", x = "Round", y = "Average Payoff") +
      facet_grid(pd_difficulty ~ horizon) +  # 2x2 design: Difficulty by Game Length
      theme_minimal() +
      theme(legend.position = "none")

    ```

#### 3.3 Average Payoff and Variance Calculation per Player

-   **Objective:** Calculate the overall average payoff and payoff variance for each player to understand the distribution of player performance.

-   **Code Updates (Nov 1, 2024) =\> 2x2 Design**

    ```{r}
    avg_payoff_variance <- df_merged %>%
      group_by(player_id, horizon, pd_difficulty) %>%
      summarize(
        avg_payoff = mean(payoff, na.rm = TRUE),
        payoff_variance = var(payoff, na.rm = TRUE),
        .groups = "drop"
      )

    # Visualization with 2x2 Design
    ggplot(avg_payoff_variance, aes(x = avg_payoff)) +
      geom_histogram(bins = 30, fill = "blue", color = "black") +
      facet_grid(pd_difficulty ~ horizon) +
      labs(title = "Distribution of Average Payoffs by Game Type", x = "Average Payoff", y = "Count") +
      theme_minimal()


    ```

#### 3.4 'T' Value Calculation and Distribution Visualization

-   **Objective:** Calculate T value (the number of remaining rounds when a player first defects) and visualize the distribution of T values.

-   **Code Updates (Nov 1, 2024) =\> 2x2 Design**

    ```{r}

    df_merged <- df_merged %>%
      group_by(player_id, supergame) %>%
      mutate(
        first_defect_round = ifelse(player_coop == 0 & !is.na(player_coop), round, NA),
        T = ifelse(!is.na(first_defect_round), horizon - first_defect_round, ifelse(all(player_coop == 1, na.rm = TRUE), 0, NA))
      ) %>%
      ungroup() 

    print(df_merged %>% select(player_id, supergame, round, player_coop, horizon, T))


    # Visualization: Distribution of T Values with 2x2 Design
    ggplot(df_merged %>% filter(!is.na(T)), aes(x = T)) +
      geom_histogram(bins = 30, fill = "green", color = "black") +
      facet_grid(pd_difficulty ~ horizon) +
      labs(title = "Distribution of T Values by Game Type", x = "T (Remaining Periods at First Defection)", y = "Count") +
      theme_minimal()
    ```

#### 3.5 Cross-Tabulation of T Values by Game Length

-   **Objective:** Summarize T values by game length (horizon) to show mean and standard deviation of T for each horizon and visualize these statistics.

-   **Code Updates (Nov 1, 2024) =\> 2x2 Design**

    ```{r}
    # Update T_summary to group by both horizon and pd_difficulty
    T_summary <- df_merged %>%
      group_by(horizon, pd_difficulty) %>%
      summarise(
        mean_T = mean(T, na.rm = TRUE),
        sd_T = sd(T, na.rm = TRUE),
        count = n(),
        .groups = "drop"
      )

    print(T_summary)

    # Visualization: Average T Value by Game Length and Difficulty
    ggplot(T_summary, aes(x = factor(horizon), y = mean_T, fill = pd_difficulty)) +
      geom_bar(stat = "identity", position = "dodge") +
      facet_wrap(~ pd_difficulty) +
      labs(title = "Average T Value by Game Length and Difficulty", x = "Game Length (Horizon)", y = "Average T") +
      theme_minimal()

    # Visualization: Standard Deviation of T Value by Game Length and Difficulty
    ggplot(T_summary, aes(x = factor(horizon), y = sd_T, fill = pd_difficulty)) +
      geom_bar(stat = "identity", position = "dodge") +
      facet_wrap(~ pd_difficulty) +
      labs(title = "Standard Deviation of T Value by Game Length and Difficulty", x = "Game Length (Horizon)", y = "Standard Deviation of T") +
      theme_minimal()


    ```

------------------------------------------------------------------------

### **Step 4: Strategy Identification**

-   **Objective:** Define a function to classify player strategies (Always Cooperate, Always Defect, Grim, Tit-for-Tat, and Experimenter) and merge the identified strategies with the main dataset.

-   **Code Updates (Nov 1, 2024) =\> 2x2 Design**

-   **Code Updates (Nov 5, 2024) =\> Added Grim strategy refinement with tolerance levels (G_0, G_1, G_2) and implemented interaction-based strategy classification.**

    ```{r}


    # 4.1: Apply Grim strategy tolerance refinement (G_0, G_1, G_2)
    df_merged <- df_merged %>%
      mutate(
        G_0 = if_else(player_coop == 1 & lag(opponent_coop, default = 1) == 0, 1, 0),
        G_1 = if_else(lag(player_coop, 1, default = 1) == 1 & lag(opponent_coop, 1, default = 1) == 0, 1, 0),
        G_2 = if_else(lag(player_coop, 2, default = 1) == 1 & lag(opponent_coop, 2, default = 1) == 0, 1, 0)
      )

    # 4.2: Define function for interaction-based strategy classification
    identify_strategy <- function(player_coop, opponent_coop, tolerance) {
      
      # Check for Always Cooperate (AC)
      if (all(player_coop == 1, na.rm = TRUE)) {
        return("Always Cooperate")
      }
      
      # Check for Always Defect (AD)
      if (all(player_coop == 0, na.rm = TRUE)) {
        return("Always Defect")
      }
      
      # Check for Grim strategy with varying levels of tolerance
      if (all((player_coop == 1) | (cumsum(opponent_coop == 0) <= tolerance), na.rm = TRUE)) {
        return(paste("Grim with Tolerance", tolerance))
      }
      
      # Check for Tit-for-Tat (TFT) - mirrors opponent's previous action
      if (length(player_coop) > 1 && all(player_coop[-1] == lag(opponent_coop)[-1], na.rm = TRUE)) {
        return("Tit for Tat")
      }
      
      # Default to Experimenters if no match
      return("Experimenters")
    }

    # 4.3: Apply the classification across interactions
    df_merged <- df_merged %>%
      group_by(player_id, opponent_id, supergame) %>%
      mutate(
        strategy_label = identify_strategy(player_coop, opponent_coop, tolerance = 0), # Strict Grim
        strategy_label_tolerance_1 = identify_strategy(player_coop, opponent_coop, tolerance = 1), # Grim T1
        strategy_label_tolerance_2 = identify_strategy(player_coop, opponent_coop, tolerance = 2) # Grim T2
      ) %>%
      ungroup()

    # 4.4: Check the consistency of strategy labels across different Grim tolerance levels
    df_merged <- df_merged %>%
      mutate(
        strategy_final = case_when(
          strategy_label == "Always Cooperate" ~ "Always Cooperate",
          strategy_label == "Always Defect" ~ "Always Defect",
          strategy_label == "Tit for Tat" ~ "Tit for Tat",
          strategy_label == "Experimenters" ~ "Experimenters",
          TRUE ~ strategy_label # Defaulting to the primary Grim strategy identified
        )
      )

    # 4.5: Print a summary of strategies by game for inspection
    strategy_summary <- df_merged %>%
      group_by(player_id, strategy_final) %>%
      summarise(
        avg_payoff = mean(payoff, na.rm = TRUE),
        payoff_variance = var(payoff, na.rm = TRUE),
        count = n(),
        .groups = "drop"
      )

    # Display strategy summary
    print(strategy_summary)

    # Visualization for average payoff by strategy
    ggplot(strategy_summary, aes(x = strategy_final, y = avg_payoff, fill = strategy_final)) +
      geom_bar(stat = "identity", position = position_dodge()) +
      labs(title = "Average Payoff by Identified Strategy", x = "Strategy", y = "Average Payoff") +
      theme_minimal()
    ```

    \

    ```{r}

    # Calculate average payoff by strategy, horizon, and PD difficulty
    strategy_payoff_difficulty <- df_merged %>%
      group_by(strategy_label, horizon, pd_difficulty) %>%
      summarise(
        avg_payoff = mean(payoff, na.rm = TRUE),
        payoff_variance = var(payoff, na.rm = TRUE),
        count = n(),
        .groups = "drop"
      )
    # Print summary table
    print(strategy_payoff_difficulty)


    #####========Updated_241101===================================================#####

    # Visualization: Average Payoff by Strategy, Horizon, and PD Difficulty with adjusted text size
    ggplot(strategy_payoff_difficulty, aes(x = strategy_label, y = avg_payoff, fill = pd_difficulty)) +
      geom_bar(stat = "identity", position = position_dodge()) +
      facet_wrap(~ horizon, scales = "free") +
      scale_fill_manual(values = c("EasyPD" = "#1f77b4", "HardPD" = "#ff7f0e", "NA" = "grey70")) +
      labs(title = "Average Payoff by Strategy, Horizon, and PD Difficulty",
           x = "Strategy",
           y = "Average Payoff",
           fill = "PD Difficulty") +
      theme_minimal() +
      theme(
        text = element_text(size = 10),  # Decrease overall text size
        axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # Rotate and reduce x-axis labels
        strip.text = element_text(size = 9),  # Decrease size of facet labels
        plot.title = element_text(size = 12, face = "bold")  # Title size adjustment
      )


    #============================================================================#
    ```

------------------------------------------------------------------------

### **Step 5: Strategy Payoff Analysis**

-   **Objective:** Analyze the average payoff and payoff variance for each identified strategy, segmented by game length (horizon) and PD difficulty level (EasyPD or HardPD). This step provides insights into the effectiveness of each strategy in different game conditions and helps identify which strategies yield higher payoffs across varying difficulty levels and game lengths.

-   **Code Updates (Nov 1, 2024) =\> 2x2 Design**

    ```{r}
    # Calculate average payoff by strategy, horizon, and PD difficulty
    strategy_payoff_summary <- df_merged %>%
      group_by(strategy_label, horizon, pd_difficulty) %>%
      summarise(
        avg_strategy_payoff = mean(payoff, na.rm = TRUE),
        payoff_variance = var(payoff, na.rm = TRUE),
        count = n(),
        .groups = "drop"
      )

    # 2x2 Design Visualization with adjusted text size
    ggplot(strategy_payoff_summary, aes(x = strategy_label, y = avg_strategy_payoff, fill = pd_difficulty)) +
      geom_bar(stat = "identity", position = position_dodge()) +
      facet_grid(pd_difficulty ~ horizon, scales = "free") +
      scale_fill_manual(values = c("EasyPD" = "#1f77b4", "HardPD" = "#ff7f0e", "NA" = "grey70")) +
      labs(title = "2x2 Design: Average Payoff by Strategy, Horizon, and PD Difficulty",
           x = "Strategy",
           y = "Average Payoff",
           fill = "PD Difficulty") +
      theme_minimal() +
      theme(
        text = element_text(size = 10),  # Adjust overall text size
        axis.text.x = element_text(angle = 45, hjust = 1, size = 8),  # Rotate and resize x-axis labels
        strip.text = element_text(size = 9),  # Decrease facet label size
        plot.title = element_text(size = 12, face = "bold")  # Adjust title size and make bold
      )


    ```
